---
title: "Temp_Survey_PCA_Analysis"
author: "Kelly Chimpen Mac Leod"
format: html
editor: visual
---


# Install packages

```{r}
#install.packages("tidyverse")
#install.packages("janitor")
#install.packages("eeptools")
#install.packages("rstatix")
```

## Load `library` packages

```{r}
library("tidyverse")
library(janitor)
library(eeptools)
library(rstatix)
```

```{r}
temp <- read.csv("~/Temp_Survey_all.csv")
```

## MEDIAN SCORES FOR ALL RHINOS

### 1. Median:

Calculate median to get 1 row per individual rhino.

```{r}
temp_median <- temp %>% 
  group_by(rhino_id) %>%
 summarize(across(where(is.numeric), ~ median(.x, na.rm = TRUE), .names = "{.col}"),
            across(where(is.character),first))
```

Note: when only 2 keepers, the median = mean.

```{r}
count(temp, rhino_id) %>% arrange(n)
```

14 rhinos have only 1 keeper. 33 rhinos have only 2 keepers.

### 2. Check missing values

```{r}
#Check what data is missing 
colSums(is.na(temp_median))
```

1 conspecific fearful and 6 preferred object are NAs.

```{r}
temp_median <- temp_median %>% 
  dplyr::select(rhino_id, inst_id, 
              species, sex, age, birth_type,
              housing, lat, long, keeper_id, social_incomp, everything()) %>% 
  dplyr::select(-keeper_id)
```

# KENDALL'S W

## Install packages:

```{r}
#install.packages(stats)
#install.packages(tidyverse)
#install.packages(reshape2)
#install.packages(rcompanion)
```

## Load `library` packages:

```{r}
library(stats)
library(tidyverse)
library(reshape2)
library(rcompanion)
```

### 1. Divide by species:

Create object for BRs by using `str_detect`and `filter` functions. Then, select all behaviors data collected for both socially and solitarily housed rhinos. Also keep rhino ID, keeper ID, and institution ID. All other columns were removed because they are not of interest for the kendall's W test.

```{r}
BR <- temp %>% filter(str_detect(species, "Black")) %>% 
  arrange(inst_id) %>% 
  dplyr::select(rhino_id, keeper_id, inst_id, active, calm, investigative, sleep, stereotypies, 
                fam_af, fam_ag, fam_ff, responsive, stranger_af, stranger_ag, stranger_ff,
                social, vocal, con_af, con_ag, con_ff, pre_obj, no_obj, food_ag, chemosensory, assertive)
```

Create object for WRs by using `str_detect`and `filter` functions. Then, select all behaviors data collected for both socially and solitarily housed rhinos. Also keep rhino ID, keeper, and institution ID.

```{r}
WR <- temp %>% filter(str_detect(species, "White")) %>% 
  arrange(inst_id) %>% 
  dplyr::select(rhino_id, keeper_id, inst_id, active, calm, investigative, sleep, stereotypies, 
                fam_af, fam_ag, fam_ff, responsive, stranger_af, stranger_ag, stranger_ff,
                social, vocal, con_af, con_ag, con_ff, pre_obj, no_obj, food_ag, chemosensory, assertive)
```

Create list of behaviors that will be used for the function below (calculate_kendallW_all). Behaviors that were asked about all rhinos:

```{r}
beh <- c("active", "calm", "investigative", "sleep", "stereotypies", "assertive", "chemosensory",
               "fam_af", "fam_ag", "fam_ff", "responsive", "food_ag", "no_obj", "pre_obj", "social", 
               "vocal", "stranger_af", "stranger_ag", "stranger_ff", "con_af", "con_ag", "con_ff")
```

### 2. Create kendall's W function:

Function to calculate Kendall's W for each rhino ID per behavior. Annotations within code chunk

```{r}
calculate_kendallW_all <- function(data, behaviors) {
  # Ensure the data is in the correct format
  data <- as.data.frame(data)
  
  # Get unique inst_ids
  rhino_ids <- unique(data$rhino_id)
  
  # Initialize a list to store results
  results <- list()
  
  # Loop through each behavior
  for (behavior in behaviors) {
    # Loop through each rhino_id
    for (rhino in rhino_ids) {
      # Subset data for the specific rhino_id and behavior
      subset_data <- data %>%
        filter(rhino_id == rhino) %>%
        dplyr::select(keeper_id, inst_id, rhino_id, !!sym(behavior)) %>%
        pivot_wider(names_from = rhino_id, values_from = !!sym(behavior)) %>%
        dplyr::select(-inst_id)
      
      # Transpose and convert to matrix (needs to be as a matrix for `kendallW` function to work)
       subset_matrix <- as.matrix((subset_data))
      
      # Remove rows with NAs
      subset_matrix <- subset_matrix[complete.cases(subset_matrix), ]

      # Check if the matrix is not empty and has the correct dimensions
      if (!is.null(dim(subset_matrix)) && ncol(subset_matrix) > 1 && nrow(subset_matrix) > 1) {
        # Calculate Kendall's W
        kendallW_result <- kendallW(subset_matrix)
        
        # Store the result
        results[[paste(rhino, behavior, sep = "_")]] <- kendallW_result
      } else {
        results[[paste(rhino, behavior, sep = "_")]] <- NA
      }
    }
  }
  
  return(results)
}
```

### 3. BR - kendall's W values:

Calculate kendall's W using function above.

```{r}
B_t <- calculate_kendallW_all(BR, beh)
```

The data won't transpose (next code chunk) unless it's a dataframe so convert from matrix to dataframe.

```{r}
B_t <- as.data.frame(B_t)
```

The kendall function created above gives me the values all together in 1 row. Transpose and keep it as a dataframe. If `as.data.frame` isn't added, it goes back to the matrix format.

```{r}
B_t <- as.data.frame(t(B_t))
```

Convert what is now the rowname to a column and name it "rhino_id".

```{r}
B_t <- rownames_to_column(B_t, var = "rhino_id")
```

The 1st column format is "rhino_id_beh". Separate into 2 columns by '-'. Remove NA rows where kendall's W wasn't calculated.

```{r}
B_t <- B_t %>%
  separate(rhino_id, into = c("rhino_id", "beh"), sep = "_", extra = "merge") %>% 
  filter(W != "NA")
```

Using `dcast`, create columns for each behavior using the beh column to reshape table.

```{r}
B_t <- dcast(B_t, rhino_id~beh)
```

Arrange behaviors: behaviors asked about all rhinos as first columns and behaviors asked about socially housed rhinos only as last columns.

```{r}
B_t <- B_t %>% dplyr::select(rhino_id, active, calm, investigative, responsive, sleep, 
                     stereotypies, fam_af, fam_ag, fam_ff, stranger_af, stranger_ag, stranger_ff, everything())
```

Capitalize first letter in column names and rename abbreviations.

```{r}
B_t <- B_t %>%
  rename_with(str_to_title) %>% 
  rename("Rhino ID" = Rhino_id) %>% 
  rename("Familiar affiliative" = Fam_af) %>% 
  rename("Familiar aggressive" = Fam_ag) %>% 
  rename("Familiar fearful" = Fam_ff) %>% 
  rename("Stranger affiliative" =Stranger_af) %>% 
  rename("Stranger aggressive" = Stranger_ag) %>% 
  rename("Stranger fearful" = Stranger_ff) %>% 
  rename("Conspecific affiliative" = Con_af) %>% 
  rename("Conspecific aggressive" = Con_ag) %>% 
  rename("Conspecific fearful" = Con_ff) %>% 
  rename("Food aggressive" = Food_ag) %>% 
  rename("Novel object" = No_obj) %>% 
  rename("Preferred object" = Pre_obj)
```

### 4. WR - kendall's W values:

Calculate kendall's W using function above.

```{r}
W_t <- calculate_kendallW_all(WR, beh)
```

The data won't transpose (next code chunk) unless it's a dataframe so convert from matrix to dataframe.

```{r}
W_t <- as.data.frame(W_t)
```

The kendall function created above gives me the values all together in 1 row. Transpose and keep it as a dataframe. If `as.data.frame` isn't added, it goes back to the matrix format.

```{r}
W_t <- as.data.frame(t(W_t))
```

Convert what is now the rowname to a column and name it "rhino_id".

```{r}
W_t <- rownames_to_column(W_t, var = "rhino_id")
```

The 1st column format is "rhino_id_beh". Separate into 2 columns by '-'. Remove NA rows where kendall's W wasn't calculated.

```{r}
W_t <- W_t %>%
  separate(rhino_id, into = c("rhino_id", "beh"), sep = "_", extra = "merge") %>% 
  filter(W != "NA")
```

Using `dcast`, create columns for each behavior using the beh column to reshape table.

```{r}
W_t <- dcast(W_t, rhino_id~beh)
```

Arrange behaviors: behaviors asked about all rhinos as first columns and behaviors asked about socially housed rhinos only as last columns.

```{r}
W_t <- W_t %>% dplyr::select(rhino_id, active, calm, investigative, responsive, sleep, 
                     stereotypies, fam_af, fam_ag, fam_ff, stranger_af, stranger_ag, stranger_ff, everything())
```

Capitalize first letter in column names.

```{r}
W_t <- W_t %>%
  rename_with(str_to_title)
```

Rename abbreviations.

```{r}
W_t <- W_t %>% 
  rename("Rhino ID"= Rhino_id) %>% 
  rename("Familiar affiliative" = Fam_af) %>% 
  rename("Familiar aggressive" = Fam_ag) %>% 
  rename("Familiar fearful" = Fam_ff) %>% 
  rename("Stranger affiliative" =Stranger_af) %>% 
  rename("Stranger aggressive" = Stranger_ag) %>% 
  rename("Stranger fearful" = Stranger_ff) %>% 
  rename("Conspecific affiliative" = Con_af) %>% 
  rename("Conspecific aggressive" = Con_ag) %>% 
  rename("Conspecific fearful" = Con_ff) %>% 
  rename("Food aggressive" = Food_ag) %>% 
  rename("Novel object" = No_obj) %>% 
  rename("Preferred object" = Pre_obj)
```

### 5. Save:

Save tables as csv files.

```{r}
#write.csv(W_t, "~/Temp_Survey_kendall_WR.csv", row.names = F)
#write.csv(B_t, "~/Temp_Survey_kendall_BR.csv", row.names = F)
```

# PCA

## Install packages.

```{r}
#install.packages("FactoMineR")
#install.packages("tidyverse")
```

## Load `library` packages.

```{r}
library(factoextra)
library(tidyverse)
```

### 1. Create dataframe with rhino info

```{r}
temp_info <- 
  temp %>% 
  dplyr::select(rhino_id, inst_id, species, dob, sex, birth_type, housing, lat, long) %>%
  distinct(rhino_id, .keep_all = T)
```

### 2. Divide by species (include all columns)

```{r}
br_id <- temp_median %>% 
  filter(str_detect(rhino_id, "Black"))
```

```{r}
wr_id <- temp_median %>% 
  filter(str_detect(rhino_id, "White"))
```

PCA runs with numeric data only so include only the behaviors that will be used for PCA.

```{r}
br_temp <- br_id %>% 
  dplyr::select('active','calm','investigative','sleep','stereotypies','fam_af','fam_ag','fam_ff',
                'responsive','stranger_af','stranger_ag','stranger_ff') 
```

```{r}
wr_temp <- wr_id %>% 
  dplyr::select('active','calm','investigative','sleep','stereotypies','fam_af','fam_ag','fam_ff',
                'responsive','stranger_af','stranger_ag','stranger_ff') 
```

# BR all Median PCA

## 1. Run PCA:

Data needs to be standardized for the PCA to run correctly.

```{r}
br_temp <- scale(br_temp)
```

Run PCA using `prcomp` function.

```{r}
br_pca <- prcomp(br_temp)
```

```{r}
?prcomp
```

Explore PCA results using `summary` function.

```{r}
summary(br_pca)
br_pca$rotation
```

Flip rotations to match PCA 1. Flip rotations to make it "activity" rather than "inactivity" for PC2

```{r}
br_pca$rotation[, 4] <- -1 * br_pca$rotation[, 4]
```

Flip x scores to match PCA 1.

```{r}
br_pca$x[, 4] <- -1 * br_pca$x[, 4]
```

## 2. Visualize PCA:

Biplot for PC1 and PC2 - shows variation of each variable and magnitude

```{r}
fviz_pca_var(br_pca, col.var = "blue", alpha.var = 1, repel = T, axes = c(1:2))
```

Look at where each individual falls on this biplot and subdivide by sex.

```{r}
fviz_pca_biplot(br_pca, col.var = "blue", 
                col.ind = br_id$sex,
                addEllipses = T,
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2))
```

Look at where each individual falls on this biplot and subdivide by housing.

```{r}
fviz_pca_biplot(br_pca, col.var = "blue", 
                col.ind = br_id$housing,
                addEllipses = T,
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2))
```

Biplot for PC3 and PC4.

```{r}
fviz_pca_var(br_pca, col.var = "blue", alpha.var = 1, repel = T, axes = c(3:4))
```

Biplot by sex.

```{r}
fviz_pca_biplot(br_pca, col.var = "black", 
                col.ind = br_id$sex,
                addEllipses = T,
                alpha.var = 1, 
                repel = T, 
                axes = c(2:3))
```

Biplot by housing.

```{r}
fviz_pca_biplot(br_pca, col.var = "black", 
                col.ind = br_id$housing,
                addEllipses = T,
                alpha.var = 1, 
                repel = T, 
                axes = c(2:3))
```

## 3. Save Biplots:

```{r}
BR_1_2_sex <- fviz_pca_biplot(br_pca, col.var = "blue", 
                col.ind = br_id$sex,
                addEllipses = T,
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2))
```

Look at where each individual falls on this biplot and subdivide by housing.

```{r}
BR_1_2_sociality <- fviz_pca_biplot(br_pca, col.var = "blue", 
                col.ind = br_id$housing,
                addEllipses = T,
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2))
```

Biplot for PC3 and PC4.

```{r}
BR_2_3 <- fviz_pca_var(br_pca, col.var = "blue", alpha.var = 1, repel = T, axes = c(2:3))
```

Biplot by sex.

```{r}
BR_2_3_sex <- fviz_pca_biplot(br_pca, col.var = "black", 
                col.ind = br_id$sex,
                addEllipses = T,
                alpha.var = 1, 
                repel = T, 
                axes = c(2:3))
```

Biplot by housing.

```{r}
BR_2_3_sociality <- fviz_pca_biplot(br_pca, col.var = "black", 
                col.ind = br_id$housing,
                addEllipses = T,
                alpha.var = 1, 
                repel = T, 
                axes = c(2:3))
```

## 4. Interpret:

Percentage contribution for multiple PCs.

```{r}
fviz_cos2(br_pca, choice = "var", axes = 1:2) #For PC1 - PC2
fviz_cos2(br_pca, choice = "var", axes = 1:3) #For PC1 - PC3
fviz_cos2(br_pca, choice = "var", axes = 1:4) #For PC1 - PC4
```

Contribution of each behavior by PC.

```{r}
fviz_contrib(br_pca, choice = "var", axes = 1)
fviz_contrib(br_pca, choice = "var", axes = 2)
fviz_contrib(br_pca, choice = "var", axes = 3)
fviz_contrib(br_pca, choice = "var", axes = 4)
```

```{r}
# Loading shows magnitude of contribution and relationship (+/-)
br_pca$rotation[,1:4] 

br_explained_variance <- br_pca$sdev^2

br_prop_variance <- br_explained_variance / sum(br_explained_variance)

br_cumulative_variance <- cumsum(br_pca$sdev^2) / sum(br_pca$sdev^2)

plot(1:length(br_cumulative_variance), br_cumulative_variance, type = "b", xlab = "Number of PCs", ylab = "Cumulative Variance Explained")
```

## 5. Choosing PCs:

Look at the eigenvalues, % of variance explained per PC

```{r}
get_eigenvalue(br_pca)
```

\% variance graphed

```{r}
fviz_eig(br_pca,
         addlabels = T,
         barfill = "darkturquoise",
         barcolor = "darkturquoise")
```

# WR all Median PCA

## 1. Run PCA:

Data needs to be standardized for the PCA to run correctly.

```{r}
wr_temp <- scale(wr_temp)
```

Run PCA using `prcomp` function.

```{r}
wr_pca <- prcomp(wr_temp)
```

Explore PCA results using `summary` function.

```{r}
summary(wr_pca)
wr_pca$rotation
```

Flip all to have (+) attributes as positive scores and vice versa

```{r}
wr_pca$rotation[, 2] <- -1 * wr_pca$rotation[, 2]
wr_pca$rotation[, 1] <- -1 * wr_pca$rotation[, 1]
wr_pca$rotation[, 3] <- -1 * wr_pca$rotation[, 3]
wr_pca$rotation[, 4] <- -1 * wr_pca$rotation[, 4]
```

Flip x scores to match BR PCA.

```{r}
wr_pca$x[, 2] <- -1 * wr_pca$x[, 2]
wr_pca$x[, 1] <- -1 * wr_pca$x[, 1]
wr_pca$x[, 3] <- -1 * wr_pca$x[, 3]
wr_pca$x[, 4] <- -1 * wr_pca$x[, 4]
```

## 2. Visualize PCA:

Biplot for PC1 and PC2 - shows variation of each variable and magnitude

```{r}
fviz_pca_var(wr_pca, col.var = "blue", alpha.var = 1, repel = T, axes = c(1:2))
```

```{r}
fviz_pca_biplot(wr_pca, col.var = "black", 
                col.ind = wr_id$sex,
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2))
```

```{r}
fviz_pca_biplot(wr_pca, col.var = "black", 
                col.ind = wr_id$housing,
                mean.point = FALSE,
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2),
                ggtheme = theme_minimal())
```

```{r}
fviz_pca_biplot(wr_pca, col.var = "blue", 
                mean.point = FALSE,
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2),
                ggtheme = theme_minimal())
```

```{r}
fviz_pca_biplot(wr_pca, col.var = "black", 
                mean.point = FALSE,
                alpha.var = 1, 
                col.ind = "blue", 
                label = "var",
                repel = T,
                axes = c(3:4),
                ggtheme = theme_minimal())
?fviz_pca_biplot
```

## 3. Save:

### 3.1 Save plot showing dimensions only:

```{r}
wr_plot1_2 <- fviz_pca_var(wr_pca, col.var = "blue", alpha.var = 1, repel = T, axes = c(1:2))
wr_plot3_4 <- fviz_pca_var(wr_pca, col.var = "blue", alpha.var = 1, repel = T, axes = c(1:2))
```

### 3.2 Save biplots:

First biplot shows difference by sex:

```{r}
WR_1_2_sociality <- fviz_pca_biplot(wr_pca, col.var = "black",
                col.ind = wr_id$housing,
                addEllipses = T,
                label = "var",
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2))
```

```{r}
WR_1_2_sex <- fviz_pca_biplot(wr_pca, col.var = "black",
                col.ind = wr_id$sex,
                addEllipses = T,
                label = "var",
                alpha.var = 1, 
                repel = T, 
                axes = c(1:2))
```

Second biplot shows all rhinos:

```{r}
WR_3_4 <- fviz_pca_biplot(wr_pca, col.var = "black", 
                mean.point = FALSE,
                 col.ind = "blue", 
                alpha.var = 1, 
                label = "var",
                repel = T, 
                axes = c(3:4),
                ggtheme = theme_minimal())
```

```{r}
WR_3_4_sex <- fviz_pca_biplot(wr_pca, col.var = "black", 
                col.ind = wr_id$sex,
                addEllipses = T,
                mean.point = FALSE,
                alpha.var = 1, 
                label = "var",
                repel = T, 
                axes = c(3:4),
                ggtheme = theme_minimal())
```

```{r}
WR_3_4_sociality <- fviz_pca_biplot(wr_pca, col.var = "black", 
                col.ind = wr_id$housing,
                addEllipses = T,
                mean.point = FALSE,
                alpha.var = 1, 
                label = "var",
                repel = T, 
                axes = c(3:4),
                ggtheme = theme_minimal())
```

```{r}
fviz_pca_var(wr_pca, col.var = "blue", alpha.var = 1, repel = T, axes = c(1,3)) 
fviz_pca_var(wr_pca, col.var = "blue", alpha.var = 1, repel = T, axes = c(1,2)) 
```

## 4. Interpret:

Percentage contribution for multiple PCs.

```{r}
fviz_cos2(wr_pca, choice = "var", axes = 1:2) #For PC1 - PC2
fviz_cos2(wr_pca, choice = "var", axes = 1:3) #For PC1 - PC3
fviz_cos2(wr_pca, choice = "var", axes = 1:4) #For PC1 - PC4
```

Contribution of each behavior by PC.

```{r}
fviz_contrib(wr_pca, choice = "var", axes = 1)
fviz_contrib(wr_pca, choice = "var", axes = 2)
fviz_contrib(wr_pca, choice = "var", axes = 3)
fviz_contrib(wr_pca, choice = "var", axes = 4)
```

```{r}
# Loading shows magnitude of contribution and relationship (+/-)
wr_pca$rotation[,1:4] 

wr_explained_variance <- wr_pca$sdev^2

wr_prop_variance <- wr_explained_variance / sum(wr_explained_variance)

wr_cumulative_variance <- cumsum(wr_pca$sdev^2) / sum(wr_pca$sdev^2)

plot(1:length(wr_cumulative_variance), wr_cumulative_variance, type = "b", xlab = "Number of PCs", ylab = "Cumulative Variance Explained")
```

## 5. Choosing PCs:

Look at the eigenvalues, % of variance explained per PC

```{r}
get_eigenvalue(wr_pca)
```

Scree plots to see the percentage of explained variables for each PC within the PCA

```{r}
fviz_eig(wr_pca,
         addlabels = T,
         barfill = "darkturquoise",
         barcolor = "darkturquoise")  
```

# PCA ROTATION TABLE

## 1. Create dataframe:

Create a table where I have behavior as the 1st column, then WRM-PC1, BRM-PC1 etc.

```{r}
wr_temp <- as.data.frame(wr_temp)
br_temp <- as.data.frame(br_temp)
rnames <- names(br_temp)
PC_table <- as.data.frame(rnames)
```

Before creating the table with behaviors and PCs, convert the rotations of each PCA as a dataframe so that I can pull the values from them easily.

```{r}
wr_rotation <- as.data.frame(wr_pca$rotation)
wr_rotation <-  dplyr::select(wr_rotation, 1:4)

br_rotation <- as.data.frame(br_pca$rotation)
br_rotation <-  dplyr::select(br_rotation, 1:4)
```

## 2. Add columns:

Now that I created a table, use `mutate` function to add PC rotation values.

```{r}
PC_table <- PC_table %>% 
  mutate(WRM_PC1 = wr_rotation$PC1) %>% 
  mutate(BRM_PC1 = br_rotation$PC1) %>% 
  
  mutate(WRM_PC2 = wr_rotation$PC2) %>% 
  mutate(BRM_PC2 = br_rotation$PC2) %>% 
  
  mutate(WRM_PC3 = wr_rotation$PC3) %>% 
  mutate(BRM_PC3 = br_rotation$PC3) %>% 
  
  mutate(WRM_PC4 = wr_rotation$PC4) %>% 
  mutate(BRM_PC4 = br_rotation$PC4) 
```

Reorder table

```{r}
PC_table <- PC_table %>%  dplyr::select(rnames,
                                  WRM_PC1,WRM_PC2,WRM_PC3,WRM_PC4,
                                  BRM_PC1,BRM_PC2,BRM_PC3,BRM_PC4)
```

# X-SCORE TABLE

## 1. Create dataframe:

Create dataframe that includes x-values for each rhino (this one doesn't include rhino name and studbook ID yet).

```{r}
BRM_x <- as.data.frame(br_pca$x[, 1:4])
```

```{r}
WRM_x <- as.data.frame(wr_pca$x[, 1:4])
```

Create column called 'row_num' for both WRM_x and wr_id dataframe to use it to join dataframes to get the rhino ID belonging to the x-scores.

```{r}
BRM_x$row_num <- seq.int(nrow(BRM_x)) 
br_id$row_num <- seq.int(nrow(br_id))
```

```{r}
WRM_x$row_num <- seq.int(nrow(WRM_x)) 
wr_id$row_num <- seq.int(nrow(wr_id))
```

## 2. Join dataframes:

Join both dataframes by 'row_num' and rearrange.

```{r}
BRM_x <- left_join(BRM_x, br_id, by= "row_num") %>%  
   dplyr::select(rhino_id, PC1, PC2, PC3, PC4) %>% 
  arrange(rhino_id)
```

```{r}
WRM_x <- left_join(WRM_x, wr_id, by= "row_num") %>%  
   dplyr::select(rhino_id, PC1, PC2, PC3, PC4) %>% 
  arrange(rhino_id)
```

## 4. Join temp with scores

```{r}
b.temp <- BRM_x %>% 
  left_join(temp_median, by = "rhino_id") %>% 
  dplyr::select(rhino_id, inst_id, species, sex, age, birth_type, housing, lat, long, social_incomp, everything())
```

```{r}
w.temp <- WRM_x %>% 
  left_join(temp_median, by = "rhino_id") %>% 
    dplyr::select(rhino_id, inst_id, species, sex, age, birth_type, housing, lat, long, social_incomp, everything())
```

```{r}
temp <- rbind(w.temp, b.temp)
```

# Install packages

```{r}
#install.packages("gridExtra")
#install.packages("ggfortify")
#install.packages("stats")
#install.packages("factoextra")
#install.packages("corrplot")
#install.packages("corrr")
#install.packages("ggcorrplot")
#install.packages("FactoMineR")
#install.packages("tidyverse")
#install.packages("broom")
#install.packages("patchwork")
#install.packages("lubridate")
#install.packages("ggiraphExtra")
#install.packages("pscl")
#install.packages("MASS")
#install.packages("car")
#install.packages("modelr")
#install.packages("rstatix")
#install.packages("FSA")
#install.packages("DHARMa")
#install.packages("emmeans")
#install.packages("pscl")
#install.packages("knitr")
#install.packages("rmarkdown")
#install.packages("htmltools")
#install.packages("superb")
#install.packages("ggeffects")
#install.packages("vegan")
```

# `Library` packages

```{r}
library(gridExtra)
library(corrplot)
library(ggplot2)
library(ggfortify)
library(stats)
library(factoextra)
library('corrr')
library(ggcorrplot)
library("FactoMineR")
library(tidyverse)
library(broom)
library(patchwork)
library(lubridate)
library(lme4)
library(lmerTest)
library(ggiraphExtra)
library(ggeffects)
library(ggpubr)
library(MASS)
library(car)
library(modelr)
library(rstatix)
library(FSA)
library(DHARMa)
library(splines)
library("emmeans")
library("pscl")
library("knitr")
library("rmarkdown")
library('htmltools')
library('superb')
library('ggeffects')
library('vegan')
```


Use clean temperament survey which has 1 row per rhino using the median: w.temp & b.temp


# PRE DATA ANALYSIS
```{r}
temp <- temp %>% 
  mutate(species = case_when(species == "Southern White" ~ "White", TRUE ~ "Black")) 
```

## Species difference

Use the `wilcox.test` function to run Mann-Whitney U test to compare traits between species

```{r}
wilcox.test(active ~ species, data = temp) # not significant
```

```{r}
wilcox.test(calm ~ species, data = temp) #significant
```

```{r}
temp %>% ggplot(aes(species,calm, color=species)) + 
  geom_boxplot()
```

```{r}
wilcox.test(investigative ~ species, data = temp) #not significant
```

```{r}
wilcox.test(sleep ~ species, data = temp) #not significant
```

```{r}
wilcox.test(stereotypies ~ species, data = temp) #significant
```

```{r}
temp %>% ggplot(aes(species,stereotypies, color=species)) + 
  geom_boxplot()
```

```{r}
wilcox.test(assertive ~ species, data = temp) #not significant
```

```{r}
wilcox.test(chemosensory ~ species, data = temp) #not significant
```

```{r}
wilcox.test(con_af ~ species, data = temp) #not significant
```

```{r}
wilcox.test(con_ag ~ species, data = temp) #not significant
```

```{r}
wilcox.test(con_ff ~ species, data = temp) #not significant
```

```{r}
wilcox.test(food_ag ~ species, data = temp) #trend
```

```{r}
wilcox.test(no_obj ~ species, data = temp) #significant
```

```{r}
temp %>% ggplot(aes(species,scale(no_obj), color=species)) + 
  geom_boxplot()
```

```{r}
wilcox.test(pre_obj ~ species, data = temp) #not significant
```

```{r}
wilcox.test(social ~ species, data = temp) #not significant
```

```{r}
wilcox.test(vocal ~ species, data = temp) #significant
```

```{r}
temp %>% ggplot(aes(species,vocal, color=species)) + 
  geom_boxplot() 
```

```{r}
wilcox.test(fam_af ~ species, data = temp) #trend
```

```{r}
wilcox.test(fam_ag ~ species, data = temp) #significant
```

```{r}
temp %>% ggplot(aes(species,fam_ag, color=species)) + 
  geom_boxplot()
```

```{r}
wilcox.test(fam_ff ~ species, data = temp) #not significant
```

```{r}
wilcox.test(responsive ~ species, data = temp) #not significant
```

```{r}
wilcox.test(stranger_af ~ species, data = temp) #significant
```

```{r}
temp %>% ggplot(aes(species,stranger_af, color=species)) + 
  geom_boxplot()
```

```{r}
wilcox.test(stranger_ag ~ species, data = temp) #significant
```

```{r}
temp %>% ggplot(aes(species,stranger_ag, color=species)) + 
  geom_boxplot()
```

```{r}
wilcox.test(stranger_ff ~ species, data = temp) #not significant
```

Calm, stereotypies, novel object, vocal, familiar aggressive, stranger affiliative, and stranger aggressive are significantly different between species.

## Save graphs:

```{r}
stereo_sp <- temp %>% ggplot(aes(species,stereotypies, color=species)) + 
  geom_boxplot() +
  theme_ggeffects() +
  labs(x='Species',
       y='Stereotypies',
       color = 'Species') +
  theme(text = element_text(family = "serif")) +
  theme(text = element_text(size = 12))
```

```{r}
fam_ag_sp <- temp %>% ggplot(aes(species,fam_ag, color=species)) + 
  geom_boxplot() +
  theme_ggeffects() +
  labs(x='Species',
       y='Familiar aggressive',
       color = 'Species') +
  theme(text = element_text(family = "serif")) +
  theme(text = element_text(size = 12))
```

```{r}
str_ag_sp <- temp %>% ggplot(aes(species,stranger_ag, color=species)) + 
  geom_boxplot() +
  theme_ggeffects() +
  labs(x='Species',
       y='Stranger aggressive',
       color = 'Species') +
  theme(text = element_text(family = "serif")) +
  theme(text = element_text(size = 12))
```

```{r}
calm_sp <- temp %>% ggplot(aes(species,calm, color=species)) + 
  geom_boxplot() +
  theme_ggeffects() +
  labs(x='Species',
       y='Calm',
       color = 'Species') +
  theme(text = element_text(family = "serif")) +
  theme(text = element_text(size = 12))
```

```{r}
str_af_sp <- temp %>% ggplot(aes(species,stranger_af, color=species)) + 
  geom_boxplot() +
  theme_ggeffects() +
  labs(x='Species',
       y='Stranger affiliative',
       color = 'Species') +
  theme(text = element_text(family = "serif"))+
  theme(text = element_text(size = 12)) 
```

```{r}
no_obj_sp <- temp %>% ggplot(aes(species,no_obj, color=species)) + 
  geom_boxplot() +
  theme_ggeffects() +
  labs(x='Species',
       y='Novel object',
       color = 'Species') +
  theme(text = element_text(family = "serif")) +
  theme(text = element_text(size = 12))
```

```{r}
vocal_sp <- temp %>% ggplot(aes(species,vocal, color=species)) + 
  geom_boxplot() +
  theme_ggeffects() +
  labs(x='Species',
       y='Vocal',
       color = 'Species') +
  theme(text = element_text(family = "serif"))+
  theme(text = element_text(size = 12))
```

```{r}
combined_sp_survey <- ((stereo_sp+ fam_ag_sp) / (str_af_sp + str_ag_sp)/
                         (calm_sp+ no_obj_sp) / (vocal_sp + plot_spacer())) +
plot_layout(guides='collect') &
  theme(legend.position='bottom')

print(combined_sp_survey)
```

Add annotation

```{r}
combined_sp_survey <- combined_sp_survey+
  plot_annotation(tag_levels = 'A')

print(combined_sp_survey)
```

```{r}
#ggsave(combined_sp_survey, filename = "~/combined_sp_survey.tiff",  height = 6, width = 10, units = "in", dpi = 400)
```

# DATA ANALYSIS

## BLACK RHINO

### Pre-data analysis

### 1. Temperament survey data:

Continuous variables - Pearson's correlation

```{r}
cor.test(b.temp$lat, b.temp$age, method = "pearson")
cor.test(b.temp$long, b.temp$age, method = "pearson")
```

There's no relationship between location and age.

Categorical variables - Chi square test

```{r}
chisq.test(b.temp$inst_id, b.temp$housing) #significant
chisq.test(b.temp$inst_id, b.temp$age) #significant
chisq.test(b.temp$inst_id, b.temp$sex) #not significant
chisq.test(b.temp$inst_id, b.temp$long) #significant
chisq.test(b.temp$inst_id, b.temp$lat) #significant
```

Significant between institution and housing. Significant between institution and longitude and latitude. However, institution and latitude and longitude are telling us the same information which is location. Thus, I'm removing both latitude and longitude.

```{r}
chisq.test(b.temp$housing, b.temp$age)
chisq.test(b.temp$housing, b.temp$sex)
chisq.test(b.temp$housing, b.temp$inst_id)
```

Housing style is significantly correlated with institution ID.

```{r}
chisq.test(b.temp$sex, b.temp$age)
chisq.test(b.temp$sex, b.temp$inst_id)
```

Not significant. Both variables were retained.

Including institution as a random effect whenever it makes a significant different after assessing it using `Anova()` to compare model with and without random effect

### 2. Combined dataset:

Check which variables should and should not be considered depending on count.

```{r}
count(b.temp,birth_type)
```

Only 4 wild rhinos out of 50. Hence, birth type wasn't assessed in BR.

```{r}
count(b.temp,sex)
```

Continuous variables - Pearson correlation coefficient.

```{r}
chisq.test(b.temp$inst_id, b.temp$age)
```

Institution and age are correlated.As mentioned above, institution will be included as a random effect in models where it makes a difference.

### 3. Scores IQR

```{r}
IQR(b.temp$PC1)
IQR(b.temp$PC2)
IQR(b.temp$PC3)
```

```{r}
summary(b.temp$PC1)
summary(b.temp$PC2)
summary(b.temp$PC3)
```

```{r}
quantile(b.temp$PC1, 0.25)
quantile(b.temp$PC1, 0.75)
```

```{r}
quantile(b.temp$PC2, 0.25)
quantile(b.temp$PC2, 0.75)
```

```{r}
quantile(b.temp$PC3, 0.25)
quantile(b.temp$PC3, 0.75)
```

# DEMO-ENVIRO FACTORS

We are keeping 3 fixed variables which are age, sex, and housing. Institution ID is added as a random effect.

### PC1:

Visualization: Graph a histogram to determine distribution.

```{r}
b.temp %>% ggplot(aes(PC1)) + geom_histogram()
```

Right skewed distribution.

```{r}
b.temp %>%
  ggplot(aes(age, PC1, color = housing)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_ggeffects()
```

No trend.

```{r}
b.temp %>%
  ggplot(aes(age, PC1, color = sex)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_ggeffects()
```

Seems like as age increases, PC1 decreases in females and increases in males. But points are all over the graph. Let's model age and sex as an interaction.

```{r}
b.temp %>%
  ggplot(aes(housing, PC1, color = age)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_ggeffects()
```

Socially housed black rhinos seem to score higher on PC1.

#### Modelling:

Including random effect.

```{r}
B1.1 <- lmer(PC1 ~ housing + age * sex + (1|inst_id), data = b.temp)
```

As interaction between sex and age.

```{r}
B1.2a <- lm(PC1 ~ sex * age + housing, data = b.temp)
```

No interaction between sex and age.

```{r}
B1.2 <- lm(PC1 ~ sex + age + housing, data = b.temp)
```

```{r}
AIC(B1.2, B1.2a) #interaction has lower AIC
```

Let's look whether random effect should be included or excluded.

```{r}
summary(B1.1) #0 variance
```

```{r}
anova(B1.1, B1.2)
anova(B1.1, B1.2a) #no difference
```

Institution explains 0 variance. Thus, shouldn't be included in the model.

```{r}
B1.3a <- lm(PC1 ~ sex * age, data = b.temp)
```

```{r}
B1.4 <- lm(PC1 ~ housing, data = b.temp)
```

```{r}
B1.5 <- lm(PC1 ~ 1, data = b.temp)
```

```{r}
AIC( B1.4, B1.2a, B1.3a, B1.5)
```

#### Diagnostic plots:

```{r}
par(mfrow = c(2,2))
plot(B1.2a)
summary(B1.2a)
```

```{r}
Anova(B1.2a)
```

Full model includes the interaction between age and sex, plus housing. None of the variables are significant and model explains only 0.2% of PC1.

```{r}
summary(B1.4)
plot(B1.4)
```

Housing does not significantly explain PC1. R\^2 is very small and p-value \> 0.05. Neither the interaction nor their individual effect significantly affect PC1. p-value \> 0.05.

```{r}
summary(B1.5)
plot(B1.5)
```

Right tail of QQ plot is off the diagonal line.

```{r}
Anova(B1.5)
```

```{r}
par(mfrow = c(1,1))
```

```{r}
simulationOutput <- simulateResiduals(B1.2a)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

```{r}
simulationOutput <- simulateResiduals(B1.4)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

```{r}
simulationOutput <- simulateResiduals(B1.5)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Even though DHARMa is mostly used for complex models, it can also be used for simple 'lm' models. (https://cran.r-project.org/web/packages/DHARMa/vignettes/DHARMa.html#lm-and-glm) shows lm is supported. In this case, I used it to supplement my QQplots to determine whether it's acceptable or not.

Null model was the best model and passed diagnostics for PC1.

### PC2:

Visualization: Graph a histogram to determine distribution.

```{r}
b.temp %>% ggplot(aes(PC2)) + geom_histogram()
```

Somewhat normal distribution.

```{r}
b.temp %>%
  ggplot(aes(sex, PC2, color = housing)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

Females seem to score higher than males.

```{r}
b.temp %>%
  ggplot(aes(age, PC2, color = housing)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

No interaction.

```{r}
b.temp %>%
  ggplot(aes(age, PC2, color = sex)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

No interaction.

#### Modelling:

```{r}
B2 <- lmer(PC2 ~ age + sex + housing + (1|inst_id), data = b.temp)
B2.1 <- lm(PC2 ~ age + sex + housing, data = b.temp)
```

Determine if random effect should be included.

```{r}
summary(B2)
anova(B2, B2.1)
```

Random effect should not be included. There's no difference between model with and without RE as shown by the anova results.

```{r}
drop1(B2.1)
```

```{r}
B2.2 <- lm(PC2 ~ sex + housing, data = b.temp)
```

```{r}
drop1(B2.2)
```

```{r}
B2.3 <- lm(PC2 ~ sex, data = b.temp)
```

```{r}
B2.4 <- lm(PC2 ~ 1, data = b.temp)
```

Choose best model.

```{r}
AIC(B2.1, B2.2, B2.3, B2.4)
```

Lowest AIC model is B2.3, which includes sex.

#### Diagnostic plots:

```{r}
par(mfrow = c(2,2))
```

```{r}
summary(B2.1)
```

```{r}
Anova(B2.1)
```

```{r}
simulationOutput <- simulateResiduals(B2.1)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

This is the full model which include age, sex and housing. Neither age not housing are significant in this model. Sex is significant. However, model as a whole is not significant p-value \> 0.05.

```{r}
summary(B2.3)
```

```{r}
Anova(B2.3)
```

```{r}
simulationOutput <- simulateResiduals(B2.3)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Sex explains 13% (R\^2 = 0.1339) of PC2. Males score 0.4992 + (-1.0191) = -0.5199, and females 0.4992 on PC2. This model has a p-value \< 0.05.

```{r}
summary(B2.2)
```

```{r}
simulationOutput <- simulateResiduals(B2.2)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

In this model with sex and housing, sex is the only significant variable. However, this model's p-value \< 0.05.

Graph simplest model:

```{r}
ggeffects::ggeffect(B2.3, terms = "sex")
```

```{r}
effects <- ggeffects::ggeffect(B2.3, terms = "sex")
```

Convert to dataframe:

```{r}
df <- as.data.frame(effects)
```

Boxplot:

```{r}
ggplot(df, aes(x=x, y=predicted)) +
  geom_boxplot() + 
  geom_errorbar(aes(ymin=predicted-std.error, ymax=predicted+std.error), width=0.2) +
  geom_point(data = b.temp, aes(sex, PC2), color = "turquoise") +
  ylab("Predicted PC2") +
  xlab("Sex") + 
  theme_ggeffects() #raw data points in turquoise
```

Thick horizontal line is prediction from B2.3 model which includes sex only. Graph includes error bars. Raw data points are shown in turquoise.

```{r}
ggplot(b.temp, aes(sex, PC2, fill=sex)) +
  geom_boxplot() + ylim(-3, 2.5) + theme_ggeffects()
```

Above is the simple boxplot looking at the raw data only.

```{r}
predictions_B2.3 <- ggpredict(B2.3, terms = "sex", add.data = F)
```

#### Graph

```{r}
ggplot(predictions_B2.3, aes(x = x, y = predicted)) +
  geom_boxplot(size = .5, width = 0.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.1)  +
  theme_ggeffects() +
    labs(x = "Sex", 
       y = "PC Score on Inactive / Active") +
  theme(text = element_text(family = "serif")) +
  showSignificance( c(1,2), 1.2, -0.05, "*") +
  ylim(-1,1.25) +
  theme(text = element_text(size = 15))
```

Save plot:

```{r}
BR_CTI2 <- ggplot(predictions_B2.3, aes(x = x, y = predicted)) +
  geom_boxplot(size = .5, width = 0.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.1)  +
  theme_ggeffects() +
    labs(x = "Sex", 
       y = "PC Score on Inactive / Active") +
  theme(text = element_text(family = "serif")) +
  ylim(-1,1.25) +
  theme(text = element_text(size = 15))
```

This plot shows only the predicted values and error bars using B2.3 model (sex).

### PC3:

Visualization:

```{r}
b.temp %>% ggplot(aes(PC3)) + geom_histogram()
```

Normally distributed.

```{r}
b.temp %>%
  ggplot(aes(sex, PC3, color = housing)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

No difference.

```{r}
b.temp %>%
  ggplot(aes(age, PC3, color = housing)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_ggeffects()
```

Interaction between housing and age.

```{r}
b.temp %>%
  ggplot(aes(age, PC3, color = sex)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

No interaction.

#### Modelling:

```{r}
B3 <- lmer(PC3 ~ sex + age * housing + (1|inst_id), data = b.temp)

B3.1 <- lm(PC3 ~ sex + age * housing , data = b.temp)
```

Determine if random effect should be included.

```{r}
par(mfrow = c(1,1))
```

```{r}
summary(B3)
```

```{r}
anova(B3, B3.1)
```

Institution ID should not be included as RE as anova shows no difference.

```{r}
drop1(B3.1)
```

```{r}
B3.2 <- lm(PC3 ~ age * housing, data = b.temp)
```

```{r}
B3.3 <- lm(PC3 ~ housing, data = b.temp)
```

```{r}
B3.4 <- lm(PC3 ~ age, data = b.temp)
```

```{r}
B3.5 <- lm(PC3 ~ 1, data = b.temp)
```

```{r}
AIC( B3.1, B3.2, B3.3, B3.4, B3.5)
```

#### Diagnostic plots:

```{r}
par(mfrow = c(2,2))
```

```{r}
summary(B3.1)
```

```{r}
simulationOutput <- simulateResiduals(B3.1)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Model passed diagnostics. However, the lines on the residuals v. predicted are not parallel.

Full model does not explain PC3 (negative R\^2).

```{r}
summary(B3.5)
```

```{r}
simulationOutput <- simulateResiduals(B3.5)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Passed diagnostics.

```{r}
summary(B3.3)
```

```{r}
simulationOutput <- simulateResiduals(B3.3)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Levene test for homogeneity of variance is significant.

```{r}
summary(B3.4)
```

```{r}
simulationOutput <- simulateResiduals(B3.4)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

```{r}
simulationOutput <- simulateResiduals(B3.2)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Passed but lines are not parallel in residual v. predicted graph.

The null model is the best model and passed diagnostics.

## WHITE RHINO

###Pre-data analysis

### 1. Temperament survey data:

Continuous variables - Pearson's correlation

```{r}
cor.test(w.temp$lat, w.temp$age, method = "pearson")
cor.test(w.temp$long, w.temp$age, method = "pearson")
```

Categorical variables - Chi square test

```{r}
chisq.test(w.temp$inst_id, w.temp$housing) #significant
chisq.test(w.temp$inst_id, w.temp$age) #significant
chisq.test(w.temp$inst_id, w.temp$sex) #not significant
chisq.test(w.temp$inst_id, w.temp$long) #significant
chisq.test(w.temp$inst_id, w.temp$lat) #significant
```

Significant between institution and housing. Significant between institution and longitude and latitude. However, institution and latitude and longitude are telling us the same information which is location. Thus, I'm removing both latitude and longitude as variables.

```{r}
chisq.test(w.temp$housing, w.temp$age)
```

Significant between housing and age. Recommendations to house females socially and males solitarily.

```{r}
chisq.test(w.temp$sex, w.temp$age)
```

Not significant.

Check which variables should and should not be considered depending on count.

```{r}
count(w.temp,birth_type) #enoguh wild individuals to include it
count(w.temp,sex) #almost twice as many females
```

Categorical variables - Chi square test

```{r}
chisq.test(w.temp$age, w.temp$sex)
chisq.test(w.temp$age, w.temp$housing) 
chisq.test(w.temp$age, w.temp$inst_id) #significant
```

Significant correlation between age and birth type. This might be coincidental so I'm keeping them as independent variables.

```{r}
chisq.test(w.temp$sex, w.temp$housing) #significant
chisq.test(w.temp$sex, w.temp$inst_id)
```

Significant correlation between sex and housing (recommended by husbandry manual).

# DEMO-ENVIRO FACTORS

### PC1:

Visualization: Graph a histogram to determine distribution.

```{r}
w.temp %>% ggplot(aes(scale(PC1))) + geom_histogram()
```

Right skewed.

```{r}
w.temp %>%
  ggplot(aes(sex, PC1, color = housing)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(age, PC1, color = housing)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(age, PC1, color = sex)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(birth_type, PC1, color = sex)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

#### Modelling:

```{r}
W1 <- lmer(PC1 ~ housing + age + sex + birth_type+ (1|inst_id), data = w.temp)

W1.2 <- lm(PC1 ~ sex + age + housing + birth_type , data = w.temp)
```

Determine if random effect should be included.

```{r}
summary(W1)
anova(W1, W1.2)
ranova(W1)
```

Include institution ID in model.

```{r}
drop1(W1)
```

Remove birth type

```{r}
W1.3 <- lmer(PC1 ~ age + sex + housing + (1|inst_id), data = w.temp)
W1.3a <- lm(PC1 ~ age + sex + housing, data = w.temp)
```

```{r}
anova(W1.3, W1.3a)
ranova(W1.3)
```

```{r}
drop1(W1.3)
```

```{r}
W1.4 <- lmer(PC1 ~ housing + sex + (1|inst_id), data = w.temp)
```

```{r}
W1.5 <- lmer(PC1 ~ housing +  (1|inst_id), data = w.temp)
```

```{r}
W1.6 <- lmer(PC1 ~ 1 + (1|inst_id), data = w.temp)
W1.6a <- lm(PC1 ~ 1, data = w.temp)
```

```{r}
AIC(W1, W1.2, W1.3, W1.3a, W1.4, W1.5, W1.6, W1.6a)
```

```{r}
anova(W1)
```

W1.5 best model with housing.

#### Diagnostic plots:

```{r}
simulationOutput <- simulateResiduals(W1.5)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

Levene test for homogeneity of variance is significant.

```{r}
simulationOutput <- simulateResiduals(W1)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

Passed diagnostics

```{r}
simulationOutput <- simulateResiduals(W1.4)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Passed diagnostics

```{r}
Anova(W1.4)
```

```{r}
summary(W1.3)
anova(W1.3)
```

Plot: best model (W1.5) didn't pass diagnostics, but next best did (W1.4).

```{r}
predictions_W1.4 <- ggpredict(W1.4, terms = c("housing","sex"))
```

```{r}
ggplot(predictions_W1.4, aes(x = x, y = predicted, color = group, fill = group)) +
 geom_boxplot(size = .5, width = 0.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.1,  
                position = position_dodge(0.5)) +
  scale_color_manual(values = c("green4", "darkviolet")) +
  scale_fill_manual(values = c("green4", "darkviolet")) +
  theme_ggeffects() +
    labs(color = "Sociality",
         fill = "Sociality",
         x = "Sex", 
       y = "Anthrophobic / Anthrophilic") +
  theme(text = element_text(family = "serif"))  +
  theme(text = element_text(size = 15))
```

```{r}
WR_CTI1 <- ggplot(predictions_W1.4, aes(x = x, y = predicted, color = group, fill = group)) +
 geom_boxplot(size = .5, width = 0.5) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), 
                width = 0.1,  
                position = position_dodge(0.5)) +
  scale_color_manual(values = c("green4", "darkviolet")) +
  scale_fill_manual(values = c("green4", "darkviolet")) +
  theme_ggeffects() +
    labs(color = "Sociality",
         fill = "Sociality",
         x = "Sex", 
       y = "Anthrophobic / Anthrophilic") +
  theme(text = element_text(family = "serif"))  +
  theme(text = element_text(size = 15))
```

### PC2:

Visualization: Graph a histogram to determine distribution.

```{r}
w.temp %>%  ggplot(aes(PC2)) + geom_histogram()
```

Looks normal except for 3 outliers.

```{r}
w.temp %>%
  ggplot(aes(sex, PC2, color = housing)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(age, PC2, color = housing)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(age, PC2, color = sex)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  theme_ggeffects()
```

#### Modelling:

```{r}
W2 <- lmer(PC2 ~ housing + age + sex + birth_type +(1|inst_id), data = w.temp)

W2.1 <- lm(PC2 ~ sex + age + birth_type+ housing, data = w.temp)
```

Determine if random effect should be included.

```{r}
summary(W2)
anova(W2, W2.1)
```

Institution should not be included as random effect.

```{r}
drop1(W2.1)
```

```{r}
W2.2 <- lm(PC2 ~ birth_type +sex + age, data = w.temp)
```

```{r}
drop1(W2.2)
```

```{r}
W2.3 <- lm(PC2 ~ sex +age, data = w.temp)
```

```{r}
drop1(W2.3)
```

```{r}
W2.4 <- lm(PC2 ~ age, data = w.temp)
```

```{r}
W2.5 <- lm(PC2 ~ 1, data = w.temp)
```

```{r}
AIC(W2.1, W2.2, W2.3, W2.4, W2.5)
```

Model with sex and age is best fit (W2.2)

#### Diagnostic plots:

```{r}
simulationOutput <- simulateResiduals(W2.1)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

Combined adjusted quantile test is significant

```{r}
summary(W2.1)
```

```{r}
anova(W2.1)
```

```{r}
simulationOutput <- simulateResiduals(W2.2)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

Passed diagnostics

```{r}
summary(W2.2)
```

```{r}
anova(W2.2)
```

```{r}
simulationOutput <- simulateResiduals(W2.3)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

```{r}
summary(W2.3)
```

Good model. Sex and age are significant. At age 0, females score 0.96 on PC2. Males score 0.495 more than females on PC2. As rhinos age, PC2 increases by 0.06 per year.

Graph models:

```{r}
plot(ggeffect(W2.2, terms = c("age", "sex"))) +
  labs(x = "Age",
       title = "Predicted: PC2")
```

#### Graph

```{r}
predictions_W2.2 <- ggpredict(W2.2, terms = c("age","sex"))
```

```{r}
ggplot(predictions_W2.2, aes(x = x, y = predicted, color = group, fill = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.1) +
  scale_color_manual(values = c("magenta", "turquoise1")) +
  scale_fill_manual(values = c("magenta", "turquoise1")) +
  theme_ggeffects() +
    labs(color = "Sex",
         fill = "Sex",
         x = "Age", 
       y = "Inactive / Active") +
 theme(text = element_text(family = "serif")) +
  theme(text = element_text(size = 15))
```

```{r}
WR_CTI2 <- ggplot(predictions_W2.2, aes(x = x, y = predicted, color = group, fill = group)) +
  geom_line(size = 1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.1) +
  scale_color_manual(values = c("magenta", "turquoise1")) +
  scale_fill_manual(values = c("magenta", "turquoise1")) +
  theme_ggeffects() +
    labs(color = "Sex",
         fill = "Sex",
         x = "Age", 
       y = "PC Score on Inactive / Active") +
 theme(text = element_text(family = "serif")) +
  theme(text = element_text(size = 15))
```

### PC3:

Visualization: Graph a histogram to determine distribution.

```{r}
w.temp %>% ggplot(aes(PC3)) + geom_histogram()
```

Normal distribution.

```{r}
w.temp %>%
  ggplot(aes(sex, PC3, color = housing)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(age, PC3, color = housing)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(age, PC3, color = sex)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

#### Modelling:

```{r}
W3 <- lmer(PC3 ~ age + sex + housing + birth_type + (1|inst_id), data = w.temp)

W3.1 <- lm(PC3 ~ age + sex + housing + birth_type , data = w.temp)
```

Determine if random effect should be included.

```{r}
summary(W3)
anova(W3, W3.1)
```

Remove institution ID as random effect.

```{r}
drop1(W3.1)
```

```{r}
W3.2 <- lm(PC3 ~ age + sex + housing  , data = w.temp)
```

```{r}
W3.3 <- lm(PC3 ~ age + housing, data = w.temp)
```

```{r}
drop1(W3.3)
```

```{r}
W3.4 <- lm(PC3 ~ age, data = w.temp)
```

```{r}
W3.5 <- lm(PC3 ~ 1, data = w.temp)
```

```{r}
AIC( W3.1, W3.2, W3.3, W3.4, W3.5)
```

The best model includes age only.

#### Diagnostic plots:

```{r}
par(mfrow = c(2,2))
```

```{r}
simulationOutput <- simulateResiduals(W3.1)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

Passed

```{r}
summary(W3.1)
```

```{r}
anova(W3.1)
```

No variable can significantly explain PC3.

```{r}
simulationOutput <- simulateResiduals(W3.5)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

Passed diagnostics.

```{r}
summary(W3.5)
```

Null model works best.

```{r}
simulationOutput <- simulateResiduals(W3.3)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

didn't pass.

```{r}
summary(W3.3)
```

```{r}
simulationOutput <- simulateResiduals(W3.2)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput)
```

```{r}
summary(W3.2)
```

Model doesn't explain PC3 either.

Best model is null model

Many raw data points are outside the line. Null model works best.

### PC4:

Visualization: Graph a histogram to determine distribution.

```{r}
w.temp %>% ggplot(aes(PC4)) + geom_histogram()
```

Normal distribution.

```{r}
w.temp %>%
  ggplot(aes(sex, PC4, color = housing)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(age, PC4, color = housing)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(housing, PC4, color = housing)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(age, PC4, color = sex)) + 
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
w.temp %>%
  ggplot(aes(birth_type, PC4, color = sex)) + 
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
```

#### Modelling:

```{r}
W4 <- lmer(PC4 ~ housing + age + sex +birth_type+ (1|inst_id), data = w.temp)

W4.1 <- lm(PC4 ~ sex + age + housing +birth_type, data = w.temp)
```

Determine if random effect should be included.

```{r}
summary(W4)
anova(W4, W4.1)
```

Include random effect.

```{r}
drop1(W4)
```

```{r}
W4.2a <- lmer(PC4 ~ sex + housing +birth_type + (1|inst_id), data = w.temp)
```

```{r}
drop1(W4.2a)
```

```{r}
W4.3a <- lmer(PC4 ~ housing +birth_type+ (1|inst_id), data = w.temp)
```

```{r}
ranova(W4.3a)
```

```{r}
drop1(W4.3a)
```

```{r}
W4.4a <- lmer(PC4 ~ birth_type+ (1|inst_id), data = w.temp)
```

```{r}
W4.5a <- lmer(PC4 ~ 1 + (1|inst_id), data = w.temp)
```

```{r}
AIC(W4, W4.2a, W4.3a, W4.4a, W4.5a)
```

W4.4a is best model, which includes birth style and random effect.

#### Diagnostic plots:

```{r}
par(mfrow = c(1,1))
```

```{r}
simulationOutput <- simulateResiduals(W4)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Passed diagnostic tests.

```{r}
summary(W4)
```

```{r}
Anova(W4)
```

No significant variable.

```{r}
simulationOutput <- simulateResiduals(W4.4a)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Model with birth type did not pass diagnostics

```{r}
simulationOutput <- simulateResiduals(W4.3a)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Did not pass Levene test.

```{r}
simulationOutput <- simulateResiduals(W4.5a)
plotQQunif(simulationOutput)
plotResiduals(simulationOutput) 
```

Passed all diagnostic tests.

```{r}
summary(W4.3a)
summary(W4.4a)
```

```{r}
anova(W4.1)
anova(W4.4a)
anova(W4.3a)
```

```{r}
predict_W2.3a <- ggpredict(W4.3a, c("housing", "birth_type"), type = 'fixed')
```

```{r}
ggplot(predict_W2.3a, aes(x, predicted, color=group, fill=group)) + geom_boxplot()
```

Null model is the best model. Thus, no variable explains PC4.

# PLOTS

For chosen models for PC2 in BR (sex) and WR (sex and age)

```{r}
BR_CTI2
```

```{r}
WR_CTI1
WR_CTI2
```

## Combine graphs

```{r}
combined_PC <- (BR_CTI2+ WR_CTI2) +
  plot_annotation(tag_levels = 'A')

print(combined_PC)
```

## Save graphs

```{r}
#ggsave(combined_PC, filename = "~/combined_PC.tiff", height = 4, width = 8, units = "in", dpi = 400)
```
